---
title: "GAM modeling"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load the data and libraries 

```{r load_data, echo = FALSE}
library(combinat)
library(mgcv)
library(RColorBrewer)
library(gridExtra)
library(ggplot2)
library(itsadug)
library(PtProcess)
library(cowplot)
library(dplyr)
library(Hmisc)
library(corrplot)
library(factoextra)
library(RColorBrewer)
#library(caret)

#setwd("C:/Users/mcgr323/projects/lyon")
setwd("~/Desktop/Dartmouth Research/Manuscripts/Hawaii")
Lyon_adj_mean <- read.csv(file="Lyon_adj_mean.csv") 

```

```{r,eval=TRUE}
#Create own theme for figures
own_theme <- theme_bw(base_size = 12) +
  theme(rect = element_blank(),
        axis.ticks = element_line(color = "black"),
        axis.text = element_text(color = "black"),
        axis.line = element_line(color = "black"),
        panel.grid.minor = element_blank())
```

#Plots of parameters by depth
```{r anova_co2, echo = FALSE}
Lyon_adj_mean$depth <- factor(Lyon_adj_mean$depth)
names(Lyon_adj_mean)

#get means by depth
depthMean<-Lyon_adj_mean%>%
  dplyr::group_by(depth) %>% 
  dplyr::summarise(across(2:14, ~mean_se(.x)))

depthMean$depth<-as.numeric(as.character(depthMean$depth))

#mean plus standard error
ggplot(depthMean, aes(x=depth, y=N$y, ymin=N$ymin, ymax=N$ymax)) + 
  geom_point(position=position_dodge(width=9), size=4, color="black") + 
  geom_errorbar(position=position_dodge(width=9), width=4)+
  xlab(expression("Depth (cm)"))+
  ylab(expression("%N"))+
  own_theme+
  expand_limits(x=0, y=0)+
  scale_x_reverse(breaks=seq(0, 100, 20))+
  coord_flip()

#boxplots
c<-ggplot(Lyon_adj_mean, aes(x=depth, y=C)) + 
  geom_jitter(shape = 16,
              size=1,
              color = "#17BEBB",
              position = position_jitter(0.21)) +
  geom_boxplot(alpha = 0) +
  xlab(expression("Depth (cm)"))+
  ylab(expression("%C"))+
  own_theme+
  theme(axis.title.y=element_blank(), axis.text.y=element_blank(), plot.margin = unit(c(0,0,0,0), "cm"))+
  expand_limits(x=0, y=0)+
  scale_x_discrete(limits = rev(levels(Lyon_adj_mean$depth)))+
  coord_flip()

n<-ggplot(Lyon_adj_mean, aes(x=depth, y=N)) + 
  geom_jitter(shape = 16,
              size=1,
              color = "#17BEBB",
              position = position_jitter(0.21)) +
  geom_boxplot(alpha = 0) +
  xlab(expression("Depth (cm)"))+
  ylab(expression("%N"))+
  own_theme+
  theme(axis.title.y=element_blank(), axis.text.y=element_blank(), plot.margin = unit(c(0,0.25,0,0), "cm"))+
  expand_limits(x=0, y=0)+
  scale_x_discrete(limits = rev(levels(Lyon_adj_mean$depth)))+
  coord_flip()

ph<-ggplot(Lyon_adj_mean, aes(x=depth, y=pH_H2O)) + 
  geom_jitter(shape = 16,
              size=1,
              color = "#17BEBB",
              position = position_jitter(0.21)) +
  geom_boxplot(alpha = 0) +
  xlab(expression("Depth (cm)"))+
  ylab(expression("pH"))+
  own_theme+
  theme(plot.margin = unit(c(0,0,0,1), "cm"))+
  scale_x_discrete(limits = rev(levels(Lyon_adj_mean$depth)))+
  coord_flip()

dph<-ggplot(Lyon_adj_mean, aes(x=depth, y=Delta_pH)) + 
  geom_jitter(shape = 16,
              size=1,
              color = "#17BEBB",
              position = position_jitter(0.21)) +
  geom_boxplot(alpha = 0) +
  xlab(expression("Depth (cm)"))+
  ylab(expression(Delta~ 'pH'))+
  own_theme+
  theme(axis.title.y=element_blank(), axis.text.y=element_blank(), plot.margin = unit(c(0,0,0,0), "cm"))+
  expand_limits(x=0, y=0)+
  scale_x_discrete(limits = rev(levels(Lyon_adj_mean$depth)))+
  coord_flip()

vwc<-ggplot(Lyon_adj_mean, aes(x=depth, y=vwc*100)) + 
  geom_jitter(shape = 16,
              size=1,
              color = "#17BEBB",
              position = position_jitter(0.21)) +
  geom_boxplot(alpha = 0) +
  xlab(expression("Depth (cm)"))+
  ylab(expression("VWC"))+
  own_theme+
  theme(axis.title.y=element_blank(), axis.text.y=element_blank(), plot.margin = unit(c(0,0,0,0), "cm"))+
  scale_x_discrete(limits = rev(levels(Lyon_adj_mean$depth)))+
  coord_flip()

dt<-ggplot(Lyon_adj_mean, aes(x=depth, y=Delta_temp)) + 
  geom_jitter(shape = 16,
              size=1,
              color = "#17BEBB",
              position = position_jitter(0.21)) +
  geom_boxplot(alpha = 0) +
  xlab(expression("Depth (cm)"))+
  ylab(expression(Delta~'Soil temperature (°C)'))+
  own_theme+
  theme(plot.margin = unit(c(0,0,0,1), "cm"))+
  scale_x_discrete(limits = rev(levels(Lyon_adj_mean$depth)))+
  coord_flip()

bac<-ggplot(Lyon_adj_mean, aes(x=depth, y=S_16S)) + 
  geom_jitter(shape = 16,
              size=1,
              color = "#17BEBB",
              position = position_jitter(0.21)) +
  geom_boxplot(alpha = 0) +
  xlab(expression("Depth (cm)"))+
  ylab(expression("16S richness"))+
  own_theme+
  theme(axis.title.y=element_blank(), axis.text.y=element_blank(), plot.margin = unit(c(0,0.25,0,0), "cm"))+
  expand_limits(x=0, y=0)+
  scale_x_discrete(limits = rev(levels(Lyon_adj_mean$depth)))+
  coord_flip()

fun<-ggplot(Lyon_adj_mean, aes(x=depth, y=S_ITS)) + 
  geom_jitter(shape = 16,
              size=1,
              color = "#17BEBB",
              position = position_jitter(0.21)) +
  geom_boxplot(alpha = 0) +
  xlab(expression("Depth (cm)"))+
  ylab(expression("ITS richness"))+
  own_theme+
  theme(axis.title.y=element_blank(), axis.text.y=element_blank(), plot.margin = unit(c(0,0,0,0), "cm"))+
  expand_limits(x=0, y=0)+
  scale_x_discrete(limits = rev(levels(Lyon_adj_mean$depth)))+
  coord_flip()

non<-ggplot(Lyon_adj_mean, aes(x=depth, y=non_crystal/1000)) + 
  geom_jitter(shape = 16,
              size=1,
              color = "#17BEBB",
              position = position_jitter(0.21)) +
  geom_boxplot(alpha = 0) +
  xlab(expression("Depth (cm)"))+
  ylab(expression('P-NCM ('~ g ~ kg^-1 ~'soil)'))+
  own_theme+
  theme(axis.title.y=element_blank(), axis.text.y=element_blank(), plot.margin = unit(c(0,0,0,0), "cm"))+
  expand_limits(x=0, y=0)+
  scale_x_discrete(limits = rev(levels(Lyon_adj_mean$depth)))+
  coord_flip()

f<-ggplot(Lyon_adj_mean, aes(x=depth, y=Fe_h_Fe_d)) + 
  geom_jitter(shape = 16,
              size=1,
              color = "#17BEBB",
              position = position_jitter(0.21)) +
  geom_boxplot(alpha = 0) +
  xlab(expression("Depth (cm)"))+
  ylab(expression('Active Fe ratio'))+
  own_theme+
  theme(plot.margin = unit(c(0,0,0,1), "cm"))+
  expand_limits(x=0, y=0)+
  scale_x_discrete(limits = rev(levels(Lyon_adj_mean$depth)))+
  coord_flip()

al<-ggplot(Lyon_adj_mean, aes(x=depth, y=Al_p/1000)) + 
  geom_jitter(shape = 16,
              size=1,
              color = "#17BEBB",
              position = position_jitter(0.21)) +
  geom_boxplot(alpha = 0) +
  xlab(expression("Depth (cm)"))+
  ylab(expression('OM Comp. Al ('~g ~ kg^-1 ~'soil)'))+
  own_theme+
  theme(axis.title.y=element_blank(), axis.text.y=element_blank(), plot.margin = unit(c(0,0,0,0), "cm"))+
  expand_limits(x=0, y=0)+
  scale_x_discrete(limits = rev(levels(Lyon_adj_mean$depth)))+
  coord_flip()

fe<-ggplot(Lyon_adj_mean, aes(x=depth, y=Fe_p/1000)) + 
  geom_jitter(shape = 16,
              size=1,
              color = "#17BEBB",
              position = position_jitter(0.21)) +
  geom_boxplot(alpha = 0) +
  xlab(expression("Depth (cm)"))+
  ylab(expression('OM Comp. Fe ('~g ~ kg^-1 ~'soil)'))+
  own_theme+
  theme(axis.title.y=element_blank(), axis.text.y=element_blank(), plot.margin = unit(c(0,0.25,0,0), "cm"))+
  expand_limits(x=0, y=0)+
  scale_x_discrete(limits = rev(levels(Lyon_adj_mean$depth)))+
  coord_flip()

plot_grid(dt, vwc, c, n, ph, dph, bac, fun, f, non, al, fe, labels = "AUTO", align="v", ncol=4, label_x = .88,
  label_y = .99)
#ggsave("Depth+properties.jpeg", width = 10, height = 8)

```

#ANOVA of parameters by depth
```{r anova_co2, echo = FALSE}
Lyon_adj_mean$depth <- factor(Lyon_adj_mean$depth)
names(Lyon_adj_mean)

bonfcorr=0.05/13 #0.003846

#get means by depth
depthMean<-Lyon_adj_mean%>%
  dplyr::group_by(depth) %>% 
  dplyr::summarise(across(2:14, ~mean_se(.x)))

depthMean$depth<-as.numeric(as.character(depthMean$depth))

#one way anova by depth
aov1<-aov(Delta_temp~depth, data = Lyon_adj_mean)
summary(aov1)

aov2<-aov(vwc~depth, data = Lyon_adj_mean)
summary(aov2)
TukeyHSD(aov2)

aov2<-aov(C~depth, data = Lyon_adj_mean)
summary(aov2)
TukeyHSD(aov2)

aov2<-aov(N~depth, data = Lyon_adj_mean)
summary(aov2)
TukeyHSD(aov2)

aov2<-aov(pH_H2O~depth, data = Lyon_adj_mean)
summary(aov2)
TukeyHSD(aov2)

aov2<-aov(Delta_pH~depth, data = Lyon_adj_mean)
summary(aov2)
TukeyHSD(aov2)

aov2<-aov(non_crystal~depth, data = Lyon_adj_mean)
summary(aov2)
TukeyHSD(aov2)

aov2<-aov(Fe_h_Fe_d~depth, data = Lyon_adj_mean)
summary(aov2)
TukeyHSD(aov2)

aov2<-aov(Al_p~depth, data = Lyon_adj_mean)
summary(aov2)
TukeyHSD(aov2)

aov2<-aov(Fe_p~depth, data = Lyon_adj_mean)
summary(aov2)
TukeyHSD(aov2)

aov2<-aov(S_16S~depth, data = Lyon_adj_mean)
summary(aov2)
TukeyHSD(aov2)

aov2<-aov(S_ITS~depth, data = Lyon_adj_mean)
summary(aov2)
TukeyHSD(aov2)


```

#One-way ANOVA and graph of CO2 by depth
```{r anova_co2, echo = FALSE}
Lyon_adj_mean$depth <- factor(Lyon_adj_mean$depth)
#get mean co2 by depth for one-way anova
df_aov <- Lyon_adj_mean %>%
  group_by(depth) %>%
  dplyr::summarise(
    count_depth = n(),
    mean_co2 = mean(co2, na.rm =TRUE),
    sd_co2 = sd(co2, na.rm = TRUE)
  )

#look at box plot of differences
CO2<-ggplot(Lyon_adj_mean, aes(x = depth, y = log(co2))) +
  geom_jitter(shape = 16,
              color = "#17BEBB",
              position = position_jitter(0.21)) +
    geom_boxplot(alpha = 0) +
  labs(
    x = expression("Depth (cm)"),
    y = expression('log [CO'[2 ]~ 'production (g C'~ m^-3 ~ hr^-1~')]'))+  
  own_theme+
  theme(legend.position = "none")+
  scale_x_discrete(limits = rev(levels(Lyon_adj_mean$depth)))+
  coord_flip()

#one way anova for co2 by depth
anova_one_way <- aov(log(co2)~depth, data = Lyon_adj_mean)
summary(anova_one_way)

#tukey post-hoc test 
TukeyHSD(anova_one_way)

#ggsave("CO2_depth_one.jpeg", width = 3.2, height = 4)
```

# Global GAM model 
```{r gam_global, echo = FALSE}
#make depth numeric
Lyon_adj_mean$depth <- as.numeric(as.character(Lyon_adj_mean$depth))
Lyon_adj_mean$Al_p <- as.numeric(as.character(Lyon_adj_mean$Al_p/1000))
Lyon_adj_mean$Fe_p <- as.numeric(as.character(Lyon_adj_mean$Fe_p/1000))
Lyon_adj_mean$non_crystal <- as.numeric(as.character(Lyon_adj_mean$non_crystal/1000))

#Global model for all depths 
M_Global = gam(log(co2) ~ s(vwc,k=4) + s(Delta_temp, k=4) + s(C, k=4) + s(N, k=4) + s(pH_H2O, k =4)+ s(Delta_pH, k=4) + s(non_crystal, k =4) + s(Fe_h_Fe_d,k=4) + s(Al_p, k=4) + s(Fe_p, k=4) + s(S_16S,k=4) + s(S_ITS, k=4) + s(depth, k =4)+ s(id, bs = "re")+ 1, data = Lyon_adj_mean, na.action = "na.fail", method ="REML")

options(scipen = 999, digits=4)

summary(M_Global)

#concurvity check
options("scipen"=0, "digits"=7)
concurvity(M_Global, full=FALSE)

#Global model for all depths without N, Fe_h_Fe_d, Al_p b/c concurvity
M_Global = gam(log(co2) ~ s(vwc,k=4) + s(Delta_temp, k=4) + s(C, k=4) + s(pH_H2O, k =4)+ s(Delta_pH, k=4) + s(non_crystal, k =4) + s(Fe_p, k=4) + s(S_16S,k=4) + s(S_ITS, k=4) + s(depth, k =4)+ s(id, bs = "re")+ 1, data = Lyon_adj_mean, na.action = "na.fail", method ="REML")

summary(M_Global)

#model check
gam.check(M_Global)

library(mgcViz)
modviz <- getViz(M_Global)

check(modviz,
      a.qq = list(method = "tnorm", 
                  a.cipoly = list(fill = "light blue")), 
      a.respoi = list(size = 0.5), 
      a.hist = list(bins = 10))

#residuals look great!


#partial residuals
# o<-plot(modviz)
# o + l_fitLine(colour = "black") + 
#     l_ciLine(mul = 2, colour = "black", linetype = 2) + 
#     l_points(shape = 19, size = 3) + 
#     own_theme +     
#     ggtitle("CO2 flux") 
```

#Plots of significant predictors from GAM (plots made by keeping the other values at their mean)
```{r gam model prediction graphs}

#gam model predictions
library(ggeffects)
#library(effects)
#pr <- ggpredict(M_Global) #if add multiple terms would facet
#plot(pr, add.data = TRUE, residuals=TRUE, residuals.line=TRUE)

try<-ggemmeans(M_Global, terms=c("vwc"), ci.lvl = 0.95) #predictions based on levels of other data to look at interactions?

mydf <- ggpredict(M_Global, terms = "vwc", back.transform = FALSE)
vwc<-ggplot() +
  geom_line(data=mydf, aes(x=x, y=predicted)) +
  geom_ribbon(data=mydf, aes(x=x, y=predicted,ymin = conf.low, ymax = conf.high), alpha = .1)+
  geom_point(data=Lyon_adj_mean, aes(x = Lyon_adj_mean$vwc, y = log(Lyon_adj_mean$co2) ,color = Lyon_adj_mean$depth))+
  scale_color_gradient(low="tan", high="blue", name = "Depth (cm)")+
                  labs(
                  x = expression('VWC'),
                  y = expression('log [CO'[2 ]~ 'production (g C'~ m^-3 ~ hr^-1~')]'))+          
                own_theme+  theme( plot.margin = unit(c(0,0,0,0), "cm"), legend.position = "none", axis.title.y=element_blank(), axis.text.y=element_blank())

try<-ggemmeans(M_Global, terms=c("Delta_temp"), ci.lvl = 0.95) #predictions based on levels of other data to look at interactions?

mydf <- ggpredict(M_Global, terms = "Delta_temp", back.transform = FALSE)
dT<-ggplot() +
  geom_line(data=mydf, aes(x=x, y=predicted)) +
  geom_ribbon(data=mydf, aes(x=x, y=predicted,ymin = conf.low, ymax = conf.high), alpha = .1)+
  geom_point(data=Lyon_adj_mean, aes(x = Lyon_adj_mean$Delta_temp, y = log(Lyon_adj_mean$co2),color = Lyon_adj_mean$depth))+
  scale_color_gradient(low="tan", high="blue", name = "Depth (cm)")+
                  labs(
                  x = expression(Delta~ 'soil temperature (°C)'),
                  y = expression('log [CO'[2 ]~ 'production (g C'~ m^-3 ~ hr^-1~')]'))+   
                #stat_smooth(method = "gam", formula = y~ s(x, k = 4), size = 1, se = TRUE, colour = "black") +
                own_theme+  theme( plot.margin = unit(c(0,0,0,0), "cm"), legend.position = "none", axis.title.y=element_blank())

mydf <- ggpredict(M_Global, terms = "Delta_pH", back.transform = FALSE)
dpH<-ggplot() +
  geom_line(data=mydf, aes(x=x, y=predicted)) +
  geom_ribbon(data=mydf, aes(x=x, y=predicted,ymin = conf.low, ymax = conf.high), alpha = .1)+
  geom_point(data=Lyon_adj_mean, aes(x = Lyon_adj_mean$Delta_pH, y = log(Lyon_adj_mean$co2) ,color = Lyon_adj_mean$depth))+
  scale_color_gradient(low="tan", high="blue", name = "Depth (cm)")+
                  labs(
                  x = expression(Delta~ 'pH'),
                  y = expression('log [CO'[2 ]~ 'production (g C'~ m^-3 ~ hr^-1~')]'))+  
                #stat_smooth(method = "gam", formula = y~ s(x, k = 4), size = 1, se = TRUE, colour = "black") +
                own_theme+  theme( plot.margin = unit(c(0,0,0,0), "cm"), legend.position = "none", axis.title.y=element_blank(), axis.text.y=element_blank())

mydf <- ggpredict(M_Global, terms = "C", back.transform = FALSE)
C<-ggplot() +
  geom_line(data=mydf, aes(x=x, y=predicted)) +
  geom_ribbon(data=mydf, aes(x=x, y=predicted,ymin = conf.low, ymax = conf.high), alpha = .1)+
  geom_point(data=Lyon_adj_mean, aes(x = Lyon_adj_mean$C, y = log(Lyon_adj_mean$co2) ,color = Lyon_adj_mean$depth))+
  scale_color_gradient(low="tan", high="blue", name = "Depth (cm)")+
                  labs(
                  x = expression('Carbon (%)'),
                  y = expression('log [CO'[2 ]~ 'production (g C'~ m^-3 ~ hr^-1~')]'))+  
                #stat_smooth(method = "gam", formula = y~ s(x, k = 4), size = 1, se = TRUE, colour = "black") +
                own_theme+  theme( plot.margin = unit(c(0,0,0,0), "cm"), legend.position = "none")

#Fix by adding more depths to prediction
mydf <- ggpredict(M_Global, terms = "depth", back.transform = FALSE)
d<-ggplot() +
  #geom_line(data=mydf, aes(x=x, y=predicted)) +
  stat_smooth(data=mydf, aes(x=x, y=predicted),method = "gam", formula = y~ s(x, k = 4), size = 1, se = FALSE, colour = "black")+
  geom_ribbon(data=mydf, aes(x=x, y=predicted,ymin = conf.low, ymax = conf.high), alpha = .1)+
  geom_point(data=Lyon_adj_mean, aes(x = Lyon_adj_mean$depth, y = log(Lyon_adj_mean$co2) ,color = Lyon_adj_mean$depth))+
  scale_color_gradient(low="tan", high="blue", name = "depth")+
                  labs(
                  x = expression('Depth (cm)'),
                  y = expression('log [CO'[2 ]~ 'production (g C'~ m^-3 ~ hr^-1~')]'))+ 
                own_theme+  theme( plot.margin = unit(c(0,0,0,0.3), "cm"),legend.position = "none",axis.title.y=element_blank())

plot_grid(dT, vwc, C, dpH, d, labels = "AUTO", align="v", ncol=2, label_x = .9,
  label_y = .98)
#ggsave("gamPredictions.jpeg", width = 8, height = 6)

```

# GAM model by each depth (20, 40, 60, 80, 100cm)
```{r gam_depth, echo =FALSE}
#load names of predictors 
d_pred <- c("C", "pH_H2O", "Delta_pH", "non_crystal", "Fe_p", 
                          "vwc", "S_16S", "S_ITS")

##run model for all depths (20, 40, 60, 80, 100cm)
#get all possible model combinations given limitations of gam
# 5 depths * 4 terms * 3 smoothers = 60 (max smoothers by gam)
pred_mat  <- combn(d_pred, 3)
allModelsList <- apply(pred_mat, 2, function(x) as.formula(
  paste0("log(co2) ~ s(Delta_temp, k=3, by = depth)+", paste("s(", x, ", k=3, by = depth)", collapse ="+")) ))

#run model for all depths (20, 40, 60, 80, 100cm)
#gather data for gam model 
d_resp <- Lyon_adj_mean[ c("co2")]
#make depth a by factor
Lyon_adj_mean$depth <- factor(Lyon_adj_mean$depth)
d_preds <- Lyon_adj_mean[c("C", "pH_H2O", "Delta_pH", "non_crystal", "Fe_p", 
                            "vwc", "Delta_temp","depth","S_16S", "S_ITS")]
data= cbind(d_resp, d_preds)

#apply the model list to get all model results for gam
allModelsResults <- lapply(allModelsList,
                           function(x) gam(x, data=data, method="REML"))

#get AIC and find max
allModelsResults_AIC <- lapply(allModelsResults,
                           function(x) AIC(x))
AIC_df <- as.data.frame(unlist(allModelsResults_AIC, recursive = TRUE, use.names = TRUE))


best <- which.min(AIC_df[,1])

#get summary of the best model 
best_model <- allModelsResults[[best]]
summary(best_model)


#model check
gam.check(best_model)

library(mgcViz)
modviz <- getViz(best_model)

check(modviz,
      a.qq = list(method = "tnorm", 
                  a.cipoly = list(fill = "light blue")), 
      a.respoi = list(size = 0.5), 
      a.hist = list(bins = 10))

#residuals look good


#partial residuals
# o<-plot(modviz)
# o + l_fitLine(colour = "black") +
#     l_ciLine(mul = 2, colour = "black", linetype = 2) +
#     l_points(shape = 19, size = 3) +
#     own_theme +
#     ggtitle("CO2 flux")
```

#Graphs with depthwise gam predictions (plots made by keeping the other values at their mean)
```{r}
#colors to match above graphs
depthScale<-scales::seq_gradient_pal("tan", "blue", "Lab")(seq(0,1,length.out=5))

Lyon_adj_mean$group<-Lyon_adj_mean$depth

#get means by depth
depthMean<-Lyon_adj_mean%>%
  dplyr::group_by(depth) %>% 
  dplyr::summarise(across(1:20, ~mean(.x)))

mydf1 <- ggpredict(best_model, terms = c("Delta_temp [-2:2, by=.1]", "depth [20]"), condition=c(non_crystal=depthMean[1,6], S_16S=depthMean[1,12], vwc=depthMean[1,18]), back.transform = FALSE)
mydf2 <- ggpredict(best_model, terms = c("Delta_temp [-1:3, by=.1]", "depth [40]"), condition=c(non_crystal=depthMean[2,6], S_16S=depthMean[2,12], vwc=depthMean[2,18]),  back.transform = FALSE)
dT_d<-ggplot() +
  geom_point(data=Lyon_adj_mean, aes(x = Delta_temp, y = log(co2) ,color = group))+
  geom_line(data=mydf1, aes(x=x, y=predicted, color=group)) +
  geom_ribbon(data=mydf1, aes(x=x, y=predicted,ymin = conf.low, ymax = conf.high, group=group), alpha = .1)+
  geom_line(data=mydf2, aes(x=x, y=predicted, color=group)) +
  geom_ribbon(data=mydf2, aes(x=x, y=predicted,ymin = conf.low, ymax = conf.high, group=group), alpha = .1)+
      facet_wrap(~group, nrow=5)+
                  scale_color_manual(values=depthScale)+
                  coord_cartesian(ylim=c(-3,2))+
                  labs(x = expression(Delta~ 'Soil Temperature (°C)'),
                  y = expression('log [CO'[2 ]~ 'production (g C'~ m^-3 ~ hr^-1~')]'))+        
                  own_theme+  theme( plot.margin = unit(c(0,0,0,0.1), "cm"), legend.position = "none", strip.text.x = element_blank())

mydf1 <- ggpredict(best_model, terms = c("non_crystal[10:28, by=1]", "depth [20]"), condition=c(Delta_temp=depthMean[1,20], S_16S=depthMean[1,12], vwc=depthMean[1,18]), back.transform = FALSE)
mydf2 <- ggpredict(best_model, terms = c("non_crystal[12:54, by=1]", "depth [40]"), condition=c(Delta_temp=depthMean[2,20], S_16S=depthMean[2,12], vwc=depthMean[2,18]), back.transform = FALSE)
mydf3 <- ggpredict(best_model, terms = c("non_crystal[15:61, by=1]", "depth [80]"), condition=c(Delta_temp=depthMean[4,20], S_16S=depthMean[4,12], vwc=depthMean[4,18]), back.transform = FALSE)
PNCM_d<-ggplot() +
  geom_point(data=Lyon_adj_mean, aes(x = Lyon_adj_mean$non_crystal, y = log(Lyon_adj_mean$co2) ,color = Lyon_adj_mean$group))+
  geom_line(data=mydf1, aes(x=x, y=predicted, color=group)) +
  geom_ribbon(data=mydf1, aes(x=x, y=predicted,ymin = conf.low, ymax = conf.high, group=group), alpha = .1)+
  geom_line(data=mydf2, aes(x=x, y=predicted, color=group)) +
  geom_ribbon(data=mydf2, aes(x=x, y=predicted,ymin = conf.low, ymax = conf.high, group=group), alpha = .1)+
  geom_line(data=mydf3, aes(x=x, y=predicted, color=group)) +
  geom_ribbon(data=mydf3, aes(x=x, y=predicted,ymin = conf.low, ymax = conf.high, group=group), alpha = .1)+
      facet_wrap(~group, nrow=5)+
                  scale_color_manual(values=depthScale)+
                  coord_cartesian(ylim=c(-3,2))+
                  labs(x = expression('P-NCM ('~ g ~ kg^-1 ~'soil)'))+        
                  own_theme+  theme( plot.margin = unit(c(0,0,0,-2), "cm"), legend.position = "none", strip.text.x = element_blank(),
                                    axis.title.y=element_blank(), axis.text.y=element_blank())

mydf1 <- ggpredict(best_model, terms = c("S_16S[75:420, by=1]", "depth [20]"), condition=c(Delta_temp=depthMean[1,20],non_crystal=depthMean[1,6], vwc=depthMean[1,18]), back.transform = FALSE)
mydf2 <- ggpredict(best_model, terms = c("S_16S[10:595, by=1]", "depth [40]"), condition=c(Delta_temp=depthMean[2,20],non_crystal=depthMean[2,6], vwc=depthMean[2,18]),  back.transform = FALSE)
mydf3 <- ggpredict(best_model, terms = c("S_16S[20:325, by=1]", "depth [80]"), condition=c(Delta_temp=depthMean[4,20],non_crystal=depthMean[4,6], vwc=depthMean[4,18]),  back.transform = FALSE)
mydf4 <- ggpredict(best_model, terms = c("S_16S[20:110, by=1]", "depth [100]"), condition=c(Delta_temp=depthMean[5,20],non_crystal=depthMean[5,6], vwc=depthMean[5,18]),  back.transform = FALSE)
S16_d<-ggplot() +
  geom_point(data=Lyon_adj_mean, aes(x = Lyon_adj_mean$S_16S, y = log(Lyon_adj_mean$co2) ,color = Lyon_adj_mean$group))+
  geom_line(data=mydf1, aes(x=x, y=predicted, color=group)) +
  geom_ribbon(data=mydf1, aes(x=x, y=predicted,ymin = conf.low, ymax = conf.high, group=group), alpha = .1)+
  geom_line(data=mydf2, aes(x=x, y=predicted, color=group)) +
  geom_ribbon(data=mydf2, aes(x=x, y=predicted,ymin = conf.low, ymax = conf.high, group=group), alpha = .1)+
  geom_line(data=mydf3, aes(x=x, y=predicted, color=group)) +
  geom_ribbon(data=mydf3, aes(x=x, y=predicted,ymin = conf.low, ymax = conf.high, group=group), alpha = .1)+
  geom_line(data=mydf4, aes(x=x, y=predicted, color=group)) +
  geom_ribbon(data=mydf4, aes(x=x, y=predicted,ymin = conf.low, ymax = conf.high, group=group), alpha = .1)+
        facet_wrap(~group, nrow=5)+
                  scale_color_manual(values=depthScale)+
                  coord_cartesian(ylim=c(-3,2))+
                  labs(x = expression('16S Richness'))+       
                  own_theme+  theme( plot.margin = unit(c(0,0,0,-2), "cm"), legend.position = "none", strip.text.x = element_blank(),
                                    axis.title.y=element_blank(), axis.text.y=element_blank())

mydf1 <- ggpredict(best_model, terms = c("vwc[.35:.445, by=.001]", "depth [40]"), condition=c(Delta_temp=depthMean[2,20], non_crystal=depthMean[2,6], S_16S=depthMean[2,12]), back.transform = FALSE)
mydf2 <- ggpredict(best_model, terms = c("vwc[.405:.455, by=.001]", "depth [60]"), condition=c(Delta_temp=depthMean[3,20], non_crystal=depthMean[3,6], S_16S=depthMean[3,12]),  back.transform = FALSE)
mydf3 <- ggpredict(best_model, terms = c("vwc[.41:.465, by=.001]", "depth [80]"), condition=c(Delta_temp=depthMean[4,20], non_crystal=depthMean[4,6], S_16S=depthMean[4,12]),  back.transform = FALSE)
mydf4 <- ggpredict(best_model, terms = c("vwc[.41:.465, by=.001]", "depth [1000]"), condition=c(Delta_temp=depthMean[4,20], non_crystal=depthMean[4,6], S_16S=depthMean[4,12]),  back.transform = FALSE)
vwc_d<-ggplot() +
  geom_point(data=Lyon_adj_mean, aes(x = Lyon_adj_mean$vwc, y = log(Lyon_adj_mean$co2) ,color = Lyon_adj_mean$group))+
  geom_line(data=mydf1, aes(x=x, y=predicted, color=group)) +
  geom_ribbon(data=mydf1, aes(x=x, y=predicted,ymin = conf.low, ymax = conf.high, group=group), alpha = .1)+
  geom_line(data=mydf2, aes(x=x, y=predicted, color=group)) +
  geom_ribbon(data=mydf2, aes(x=x, y=predicted,ymin = conf.low, ymax = conf.high, group=group), alpha = .1)+
  geom_line(data=mydf3, aes(x=x, y=predicted, color=group)) +
  geom_ribbon(data=mydf3, aes(x=x, y=predicted,ymin = conf.low, ymax = conf.high, group=group), alpha = .1)+
        facet_wrap(~group, nrow=5)+
                  scale_color_manual(values=depthScale)+
                  coord_cartesian(ylim=c(-3,2))+
                  labs( x = expression('Volumetric Water Content'))+       
                 own_theme+  theme( plot.margin = unit(c(0,0.3,0,-2), "cm"), legend.position = "none", strip.text.x = element_blank(),
                                    axis.title.y=element_blank(), axis.text.y=element_blank())


plot_grid(dT_d, NULL, vwc_d, NULL, PNCM_d, NULL, S16_d, align="hv", rel_widths=c(1,-.1,1,-.1,1,-.1,1), nrow=1, labels = c("A"," ","B"," ", "C"," ", "D"),label_x = .8,
  label_y = 1)
#ggsave("gamDepthPredictions_REML.jpeg", width = 10, height = 8)


```

#Predictions of soil respiration response due to warming based on depth-wise GAM
```{r}
predicts_20 <- ggpredict(best_model, terms = c("Delta_temp [0,2]", "depth [20]"), condition=c(non_crystal=depthMean[1,6], S_16S=depthMean[1,12], vwc=depthMean[1,18]), back.transform = TRUE)
predicts_40 <- ggpredict(best_model, terms = c("Delta_temp [0,2]", "depth [40]"), condition=c(non_crystal=depthMean[2,6], S_16S=depthMean[2,12], vwc=depthMean[2,18]), back.transform = TRUE)

```

#Carbon relationships and miscellaneous stats
```{r}

Lyon_adj_mean<-Lyon_adj_mean%>%
  mutate(CperPNCM=(C*10)/non_crystal)%>%
  mutate(PNCMperC=non_crystal/(C*10))

#get means by depth
depthMean<-Lyon_adj_mean%>%
  dplyr::group_by(depth) %>% 
  dplyr::summarise_at(c("CperPNCM","PNCMperC"), ~mean_se(.x))

#look at box plot of differences
NCperC<-ggplot(Lyon_adj_mean, aes(x = depth, y = PNCMperC)) +
  geom_jitter(shape = 16,
              color = "#17BEBB",
              position = position_jitter(0.21)) +
    geom_boxplot(alpha = 0) +
  labs(
    x = expression("Depth (cm)"),
    y = expression('g P-NCM/g C'))+  
  scale_y_continuous(limits=c(0,6))+
  own_theme+
  theme(legend.position = "none")+
  scale_x_discrete(limits = rev(levels(Lyon_adj_mean$depth)))+
  coord_flip()

#one way anova for co2 by depth
anova_one_way <- aov(PNCMperC~depth, data = Lyon_adj_mean)
summary(anova_one_way)

#tukey post-hoc test 
TukeyHSD(anova_one_way)

#ggsave("CperPNCM_depth.jpeg", width = 3, height = 4)

#misc stats
ggplot() +
  geom_point(data=Lyon_adj_mean, aes(x = C, y = S_16S, color = depth))+
  geom_smooth(data=Lyon_adj_mean, aes(y=S_16S, x=C), method="lm", color="black")+
 labs(
    x = expression("Carbon (%)"),
    y = expression('P-NCM ('~ g ~ kg^-1 ~'soil)'))+
   scale_color_manual(values=depthScale)+
  own_theme

summary(lm(data=Lyon_adj_mean, S_16S~C))

ggplot() +
  geom_point(data=Lyon_adj_mean, aes(x = vwc, y = Delta_temp, color = depth))+
  geom_smooth(data=Lyon_adj_mean, aes(x = vwc, y = Delta_temp), method="lm", color="black")+
 labs(
    x = expression("Carbon (%)"),
    y = expression('P-NCM ('~ g ~ kg^-1 ~'soil)'))+
   scale_color_manual(values=depthScale)+
  own_theme

summary(lm(data=Lyon_adj_mean, vwc~Delta_temp))
```

#PCA matrix and plot
```{r}
df <- Lyon_adj_mean[,c(2:10,12,14,17,18,20)]

# df <- df %>% 
#   rename(
#     P_NCM = non_crystal,
#     Active_Fe_Ratio = Fe_h_Fe_d,
#     OM_Comp_Al = Al_p,
#     OM_Comp_Fe = Fe_p
#     )

options(digits=3) #just so we don't get so many digits in our results

p<-ncol(df)#no of variables
R<-cor(df)#saving the correlation matrix
e<-eigen(R) #solving for the eigenvalues and eigenvectors from the correlation matrix
L<-e$values #placing the eigenvalues in L
Vm<-matrix(0,nrow=p,ncol=p) #creating a p x p matrix with zeroes.
#Vm is an orthogonal matrix since all correlations between variable are 0.
diag(Vm)<-L #putting the eigenvalues in the diagonals
comp.matrix<-e$vectors %*% sqrt(Vm) #sometimes referred to as P matrix
#or eigenvectors x sqrt(Vm): P %*% t(P) is equal to the R matrix.
comp.matrix
library(psych)
pca<-principal(df,nfactor=p,rotate="none") #forcing to extract p=6 components
pca
loadings<-e$vectors %*% sqrt(Vm) #these are the correlations as we have shown earlier
loadings #signs are just different
L/p

res.pca <- prcomp(df, scale = TRUE, center = TRUE)

#with variable arrows
scale_color_gradient(low="tan", high="blue",5)
p <- fviz_pca_biplot(res.pca, repel = TRUE,
                habillage = Lyon_adj_mean$depth,  # Individuals color
                label = "var", 
                addEllipses=TRUE,ellipse.typ="confidence",
                col.var = "grey45", # Variables color 
               ggtheme = theme_minimal(base_size = 18), title = "")

colfunc <- colorRampPalette(c("tan", "blue"))
p + scale_color_manual(values=c(colfunc(5)))+ theme_minimal(base_size = 18)+  theme(legend.position="top")

#without variable arrows
p <- fviz_pca_ind(res.pca,
             geom = "point",
             habillage = Lyon_adj_mean$depth,  # Individuals color
             fill.ind = as.factor(Lyon_adj_mean$depth),
             addEllipses = TRUE, ellipse.type="confidence", # confidence ellipses
             mean.point = FALSE,
             legend.title = "Depth (cm)")
colfunc <- colorRampPalette(c("tan", "blue"))
p + scale_color_manual(values=c(colfunc(5)))+ scale_fill_manual(values=c(colfunc(5)))+own_theme+  
   labs(x = expression("PC1 (42.6%)"), y = expression("PC2 (16.8%)"))+
  theme(legend.position="top", plot.title=element_blank())

ggsave("PCA.jpeg", width = 5, height = 4)
                       
```





